Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-07-25 demystify.tex
updates

Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}. These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process. The size and the review load of the adcom depend on the department size. At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications. Note that most large schools, including Mason, have separate adcoms for MS programs.

Each application is assigned to about 3 faculty members, who will evaluate your profile and reach a consensus.  While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.

@@

For example, I only review those that are assigned to me, which is typically around 50--70 applications for the Fall admission. 
I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam). However, even if I look at them, I usually do not get involved in their evaluation.

@@

\paragraph{Core CS areas} Common CS areas that you should know are:
\begin{itemize}
  \item Programming Foundation (e.g., programming in C/C++ or Java, OOP)
  \item Discrete Math (e.g., logic, set theory, proof techniques)
  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
  \item Computer OS or Systems (e.g., memory management, file systems, processes)
\end{itemize}

>>>>

Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}. These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process. The size and the review load of the adcom depend on the department size. At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review about 30 applications. Note that most large schools, including Mason, have separate adcoms for MS programs.

Each application is assigned to about 3 adcom members, who will evaluate your profile and reach a consensus.  While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.

@@

For example, I only review those that are assigned to me.
I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam). However, even if I look at them, I usually do not get involved in their evaluation.

@@

\paragraph{Core CS topics} Common CS knowledge that you should know are:
\begin{itemize}
  \item Programming Foundation (e.g., programming concepts in some modern languages)
  \item Discrete Math (e.g., logic, set theory, proof techniques)
  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
  \item Computer OS or Systems (e.g., memory management, file systems, processes)
\end{itemize}
